\bf Faster Neural Machine Translation Inference

### 2018-02-02 fast-beam/fast-beam.tex
fast beam paper

As neural machine translation (NMT) models have become the new state-of-the-art, the challenge is to make their deployment efficient and economical. We have identified two areas where typical NMT models differ significantly from other deep-learning models that impact on their inference speed. Firstly, NMT models usualy contains a large number of output classes, corresponding to the output vocabulary. Secondly, rather than a single label, the output from machine translation models is a sequence of class labels that makes up the words or subwords in the target sentence. The sentence lengths are unpredictable, leading to inefficiencies during batch processing. We provide solutions for each of these issues which can increase batched inference speed by up to 41\% on modern GPUs without affecting model quality.

@@

Firstly, the number of classes in many deep-learning applications is small. However, the number of classes in NMT models is typically in the tens or hundreds of thousands, corresponding to the vocabulary size of the output language. For example, ~\cite{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary sizes of 60,000 and 90,000 subword units. This makes the output layer of NMT models very computationally expensive. Figure~\ref{fig:pie-time-europarl} shows the breakdown of amount of time during translation our NMT system; nearly 70\% of the time is involved in the output layer. We will look at optimizations which explicitly target the output layer.
\begin{figure}
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.3]{pie-time-europarl.png}} 
\end{tabular}
\caption{Proportion of time spent during translation (Europarl test set)}
\label{fig:pie-time-europarl}
\end{figure}

Secondly, the use of mini-batching is critical for fast model inference. However, mini-batching does not take into acccount variable sentence lengths which decrease the actual number of input or output sentences that are processed in parallel, negating the benefit of mini-batching. This can be partially reduced in the encoding with \emph{maxi-batching}, i.e. pre-sorting sentences by length before creating mini-batches with similar length source sentences. However, maxi-batching can introduce unacceptable delay in response for online applications as the input are not processed in the order they came. It is also not appropriate for applications where only the output lengths varies, for example, image captioning.

@@

\begin{figure}
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.3]{batch-size.png}} 
\end{tabular}
\caption{Actual batchsize during decoding (Europarl test set)}
\label{fig:batch-size}
\end{figure}

@@

There has been little work on batching algorithms, which is suprising considering its critical importance for fast model inference. ~\cite{Neubig-autobatching} describe an novel batching algorithm, however, its aim is to alleviate the burden on developers in batching of new models, not faster batching.

@@

We will focus on decoding as this is the more compute-intensive step, and issues with differing sentence sizes in encoding can partly be ameriorated by maxi-batching.

@@

We trained a sequence-to-sequence, encoder-decoder NMT system similar to that described in ~\cite{sennrich-haddow-birch:2016:P16-12}. This uses recurrent neural networks with gated recurrent units. The input and output vocabulary size were both set to 85,000 subwords using byte-pair encoding (BPE) to adjust the vocabulary to the desired size. The hidden layer dimensions was set to 512. %We used the Marian toolkit to train our models.

@@

\begin{table}
\begin{center}
\small
\begin{tabular}{|l|r|r|} \hline
		  & Europarl		& OpenSubtitles \\ \hline
\# sentences  	  & 30,000 		& 50,000 \\
\# subwords 	  & 787,908 		& 467,654 \\ 
Avg subwords/sent & 26.3		& 9.4 \\
Std dev subwords/sent & 14.9		& 6.1 \\ \hline
\end{tabular}
\end{center}
\caption{Test sets}
\label{tab:corpora}
\end{table}

@@

Fusing the softmax and beam search increase the speed of the output layer by 43\% for beam size 1, decreasing to 25\% for a beam of 10 when translating the Europarl dataset. This led to an overall increase in translation speed of up to 23\%, Figure~\ref{fig:beam-europarl}. Translation speed improved by up to 41\% when translating the Opensubtitles datset, Figure~\ref{fig:beam-opensubtitles}. Compared to translating the Europarl test st, Figure~\ref{fig:pie-time-europarl} in Section~\ref{sec:Introduction}, the amount of time taken by the output layer dominates translation time for the OpenSubtitles test set, Figure~\ref{fig:pie-time-opensubtitles}.

@@

\begin{figure}
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.3]{batch-size-small.png}} 
\end{tabular}
\caption{Actual batchsize during decoding}
\label{fig:batch-size-small}
\end{figure}

>>>>

As neural machine translation (NMT) models have become the new state-of-the-art, the challenge is to make their deployment efficient and economical. We have identified two areas where typical NMT models differ significantly from other deep-learning models that impact on their inference speed. Firstly, NMT models usually contains a large number of output classes, corresponding to the output vocabulary. Secondly, rather than a single label, the output from machine translation models is a sequence of class labels that makes up the words or sub-words in the target sentence. The sentence lengths are unpredictable, leading to inefficiencies during batch processing. We provide solutions for each of these issues which can increase batched inference speed by up to 41\% on modern GPUs without affecting model quality.

@@

Firstly, the number of classes in many deep-learning applications is small. However, the number of classes in NMT models is typically in the tens or hundreds of thousands, corresponding to the vocabulary size of the output language. For example, ~\cite{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary sizes of 60,000 and 90,000 sub-word units. This makes the output layer of NMT models very computationally expensive. Figure~\ref{fig:pie-time-europarl} shows the breakdown of amount of time during translation our NMT system; nearly 70\% of the time is involved in the output layer. We will look at optimizations which explicitly target the output layer.
\begin{figure}
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.3]{pie-time-europarl.png}} 
\end{tabular}
\caption{Proportion of time spent during translation (Europarl test set)}
\label{fig:pie-time-europarl}
\end{figure}

Secondly, the use of mini-batching is critical for fast model inference. However, mini-batching does not take into account variable sentence lengths which decrease the actual number of input or output sentences that are processed in parallel, negating the benefit of mini-batching. This can be partially reduced in the encoding with \emph{maxi-batching}, i.e. pre-sorting sentences by length before creating mini-batches with similar length source sentences. However, maxi-batching can introduce unacceptable delay in response for online applications as the input are not processed in the order they came. It is also not appropriate for applications where only the output lengths varies, for example, image captioning.

@@

\begin{figure}
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.3]{batch-size.png}} 
\end{tabular}
\caption{Actual batch size during decoding (Europarl test set)}
\label{fig:batch-size}
\end{figure}

@@

There has been little work on batching algorithms, which is surprising considering its critical importance for fast model inference. ~\cite{Neubig-autobatching} describe an novel batching algorithm, however, its aim is to alleviate the burden on developers in batching of new models, not faster batching.

@@

We will focus on decoding as this is the more compute-intensive step, and issues with differing sentence sizes in encoding can partly be ameliorated by maxi-batching.

@@

We trained a sequence-to-sequence, encoder-decoder NMT system similar to that described in ~\cite{sennrich-haddow-birch:2016:P16-12}. This uses recurrent neural networks with gated recurrent units. The input and output vocabulary size were both set to 85,000 sub-words using byte-pair encoding (BPE) to adjust the vocabulary to the desired size. The hidden layer dimensions was set to 512. %We used the Marian toolkit to train our models.

@@

\begin{table}
\begin{center}
\small
\begin{tabular}{|l|r|r|} \hline
		  & Europarl		& OpenSubtitles \\ \hline
\# sentences  	  & 30,000 		& 50,000 \\
\# sub-words 	  & 787,908 		& 467,654 \\ 
Avg sub-words/sent & 26.3		& 9.4 \\
Std dev subwords/sent & 14.9		& 6.1 \\ \hline
\end{tabular}
\end{center}
\caption{Test sets}
\label{tab:corpora}
\end{table}

@@

Fusing the softmax and beam search increase the speed of the output layer by 43\% for beam size 1, decreasing to 25\% for a beam of 10 when translating the Europarl dataset. This led to an overall increase in translation speed of up to 23\%, Figure~\ref{fig:beam-europarl}. Translation speed improved by up to 41\% when translating the Opensubtitles dataset, Figure~\ref{fig:beam-opensubtitles}. Compared to translating the Europarl test st, Figure~\ref{fig:pie-time-europarl} in Section~\ref{sec:Introduction}, the amount of time taken by the output layer dominates translation time for the OpenSubtitles test set, Figure~\ref{fig:pie-time-opensubtitles}.

@@

\begin{figure}
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.3]{batch-size-small.png}} 
\end{tabular}
\caption{Actual batch size during decoding}
\label{fig:batch-size-small}
\end{figure}
